### 2023-11-26 report/paper.tex
add initial references

\documentclass[12pt]{article}
\usepackage[legalpaper, portrait, margin=1in]{geometry}
\usepackage{multicol}
\usepackage[english]{babel}

\usepackage{xcolor}
\usepackage{blindtext}

\title{Automatic Lyric Transcription}
\author{Victor Shan}
\begin{document}
\color{red}
\maketitle
\begin{multicols*}{2}
\begin{abstract}
\blindtext
\end{abstract}

\section*{Scholarly Articles}
\begin{enumerate}
    \item \cite{wav2vecTransfer}
    \item \cite{wav2vec}
    \item \cite{Librispeech}
    \item \cite{whisper}
    \item \cite{LyricAlignAcoustic}
    \item \cite{Audio2Score}
    \item \cite{PhonemeInformed}
    \item \cite{Hansen}
    \item \cite{KaraokeApp}
    \item \cite{BayesianAST}
    \item \cite{whisperX}
    \item \cite{CSD}
    \item \cite{DALI}
    \item \cite{NUSDataset}
    \item \cite{JamendoLyrics}
    \item \cite{musDB18}
    \item \cite{Demux}
    \item \cite{CMUDict}
    \item \cite{SpecAugment}
    \item \cite{PhonationMode}
\end{enumerate}
\section*{Other Sources}
\begin{enumerate}
    \item \cite{MFCC}
    \item \cite{CTC}
\end{enumerate}
\bibliographystyle{apalike}
\bibliography{references}
\end{multicols*}
\end{document}