### 2023-11-26 report/paper.tex
add references and basic outline

\documentclass[12pt]{article}
\usepackage[legalpaper, portrait, margin=1in]{geometry}
\usepackage{multicol}
\usepackage[english]{babel}

\usepackage{xcolor}
\usepackage{blindtext}

% Paper requirement 1.5 line spacing. Remove for github publishing
\usepackage{setspace}
\onehalfspacing

\title{Automatic Lyric Transcription}
\author{Victor Shan}
\begin{document}
\color{red}
\maketitle
\begin{multicols*}{2}
\begin{abstract}
\blindtext
\end{abstract}

\section{Introduction}
\begin{itemize}
    \item What is Automatic Lyric Transcription?
    \item Why is it important?
    \item What are the challenges? (Why can't we use ASR?)
    \item What are the applications?
    \item What is the expected output?
\end{itemize}

\section{Datasets}
\begin{itemize}
    \item What are the requirements for a dataset?
    \item Why are there so few datasets?
    \item Challenges of creating new datasets?
    \item What are the existing datasets?
\end{itemize}
\subsection{Dataset Augmentation}
\subsubsection{SpecAugment}
\subsubsection{Transforming Existing Datasets}
\subsubsection{Student-Teacher Approach}

\subsection{Existing Datasets}
\subsubsection{TIMIT}
\subsubsection{LibriSpeech}
\subsubsection{JamendoLyrics}
Remember to record the performance of the state of the art models on this dataset
\subsubsection{Children's Songs Dataset}
\subsubsection{MUSDB18}
\subsubsection{NUS Dataset}
\subsubsection{Other Datasets}
\begin{itemize}
    \item DALI
    \item DAMP! (Stanford - Smule dataset)
    \item
\end{itemize}

\section{Related Works}
\subsection{Early HMM-Based Acoustic Models}
\subsection{Music Informed Models}
\subsection{wav2vec 2.0 and Transfer Learning}
\subsection{Whisper Word-Level Alignment}

\section{Method}

\section{Results}

\subsection{WER}

\subsection{PER}

\section{Discussion}

\section{Future Work}
\subsection{Adversarial Training}

\section{Conclusion}

\section*{Scholarly Articles}
\begin{enumerate}
    \item \cite{wav2vecTransfer}
    \item \cite{wav2vec}
    \item \cite{Librispeech}
    \item \cite{whisper}
    \item \cite{LyricAlignAcoustic}
    \item \cite{Audio2Score}
    \item \cite{PhonemeInformed}
    \item \cite{Hansen}
    \item \cite{KaraokeApp}
    \item \cite{BayesianAST}
    \item \cite{whisperX}
    \item \cite{CSD}
    \item \cite{DALI}
    \item \cite{NUSDataset}
    \item \cite{JamendoLyrics}
    \item \cite{musDB18}
    \item \cite{Demux}
    \item \cite{CMUDict}
    \item \cite{SpecAugment}
    \item \cite{PhonationMode}
\end{enumerate}
\section*{Other Sources}
\begin{enumerate}
    \item \cite{MFCC}
    \item \cite{CTC}
\end{enumerate}
\bibliographystyle{apalike}
\bibliography{references}
\end{multicols*}
\end{document}